\textsf{jaxoplanet}: Hardware-Accelerated Computation of Stellar Light Curves

### 2024-10-15 latex/paper.tex
feat: realistic speed

\subsection{Kepler's equation solver}
\todo{}

@@

\subsection{Precision}

@@

One of the advantage of the \textsf{JAX} library is its ability to perform hardware-accelerated computations on CPUs and GPUs. In this section, we evaluate the speed of the \textsf{jaxoplanet} implementation and compare it with the C++ implementation of \textsf{starry}. For quadratically limb darkened light curves, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the time required to compute the occultation light curve of a limb darkened star and that of a non-uniform star described by spherical harmonics with a maximum degrees $l_{max}=20$, depending on the order $q$ of the numerical integration. On CPU, our benchmark is done on the single core of Apple M2 max chip, while on GPU we use the single core of an NVIDIA Tesla V100 processor.\\\\
\begin{figure}[H]
    \begin{center}
        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_order.pdf}
        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l_{max}=20$ (right) depending on the order $q$ of the Gauss-Legendre quadrature used to compute the $\mathcal{P}$ integral numerically (see \autoref{solution_vector}). The processing time reported for \textsf{exoplanet} and \textsf{starry} is independent of $q$ as these two codes provide closed-form solutions for the integral $\mathcal{P}$. Light curves are computed for $10\,000$ points in transit and an occultor radius $r=0.1$.}
        \label{fig:speed_order_gausslegendre}
    \end{center}
\end{figure}
\begin{figure}[H]
    \begin{center}
        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N.pdf}
        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l_{max}=20$ (right) depending on the number of points in the light curve. Light curves are computed at $q=200$ and an occultor radius $r=0.1$.}
        \label{fig:speed_N}
    \end{center}
\end{figure}
\autoref{fig:speed_order_gausslegendre}, in combination with \autoref{fig:precision_order_gausslegendre_degree}, can be used to determine a value of $n$ balancing precision and computation time.\\\\
Finally, \autoref{fig:speed_N} shows the evaluation time of \textsf{jaxoplanet} as a function of the number of points in the light curve, compared to \textsf{starry} for a surface map of degree $l_{max} = 20$, and \textsf{exoplanet} for a quadratically limb darkened star. As in the previous figure, we show that \textsf{jaxoplanet} is competitive against state-of-the-art codes on CPUs and offers a clear advantage on GPUs for large datasets.

\section{Discussion}
\todo{}

>>>>

\subsection{Kepler's equation solver}
\todo{}
\subsection{Light curve functions and transformations}
\todo{}

@@

\subsection{Precision}\label{precision}

@@

One of the advantage of the \textsf{JAX} library is its ability to perform hardware-accelerated computations on CPUs and GPUs. In this section, we evaluate the speed of the \textsf{jaxoplanet} implementation and compare it with the C++ implementation of \textsf{starry}. For quadratically limb darkened light curves, we also compare our implementation with the one provided by the \textsf{exoplanet} Python package.\\\\
In \autoref{fig:speed_order_gausslegendre}, we show the evolution of the time required to compute the occultation light curve of a limb darkened star and that of a non-uniform star described by spherical harmonics of degree $l_{max}=20$, depending on the order $q$ of the numerical integration. On CPU, our benchmark is done on the single core of an Apple M2 max chip, while on GPU we use the single core of an NVIDIA Tesla V100 processor.\\\\
\begin{figure}[H]
    \begin{center}
        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_order.pdf}
        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l_{max}=20$ (right) depending on the order $q$ of the Gauss-Legendre quadrature used to compute the $\mathcal{P}$ integral numerically (see \autoref{solution_vector}). The processing time reported for \textsf{exoplanet} and \textsf{starry} is independent of $q$ as these two codes provide closed-form solutions for the integral $\mathcal{P}$. Light curves are computed for $10\,000$ points in transit and an occultor radius $r=0.1$.}
        \label{fig:speed_order_gausslegendre}
    \end{center}
\end{figure}
Finally, \autoref{fig:speed_N} shows the evaluation time of \textsf{jaxoplanet} as a function of the number of points in the light curve compared to \textsf{exoplanet}, for a quadratically limb darkened star, and \textsf{starry}, for a surface map of degree $l_{max} = 20$.
\begin{figure}[H]
    \begin{center}
        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N.pdf}
        \caption{Computation time of the occultation light curve of a limb-darkened star (left) and a non-uniform star described by spherical harmonics up to order $l_{max}=20$ (right) depending on the number of points in the light curve. Light curves are computed at $q=200$ and an occultor radius $r=0.1$.}
        \label{fig:speed_N}
    \end{center}
\end{figure}
These benchmarks are produced for values of $q$ which guarantee that our models reach the same precision as state-of-the-art codes. As discussed in \autoref{discussion}, \textsf{jaxoplanet} can be further optimized by reducing the order $q$ of numerical integration while keeping the precision required for specific applications.

\section{Discussion}\label{discussion}

In \autoref{precision}, we showed that \textsf{jaxoplanet} is on par with the C++ implementation of \textsf{starry} in terms of precision. As relatively high values of $q$ are required to achieve this performance, results from \autoref{speed} are deceiving and suggest that \textsf{jaxoplanet} is about an order of magnitude slower than \textsf{exoplanet} or \textsf{starry} on CPU, and only advantageous on GPU for very large datasets. Although we expect to further optimize our \textsf{JAX} implementation and bridge this gap in the future, \textsf{jaxoplanet} can currently outperform state-of-the-art codes for most realistic datasets. Indeed, models computed at low values of $q$, which are very fast, match the precision of measurements from modern instrumentation.\\\\
To demonstrate our point, we first consider the transmission spectroscopy of TRAPPIST-1 b performed with the JWST/NIRISS instrument (GO-2589, \citealt{Lim2023}). These observations reached an estimated precision of 89 ppm for a planet with a relative radius of the order $r\sim0.1$. In this context, numerical integration with an order as low as $q=10$ guarantees that our model reaches a precision of $10^{-13}$, an execution time on par with \textsf{exoplanet} on CPU and up to two orders of magnitude faster on GPU for large datasets (see left plot of \autoref{real_benchmark}). The second example we consider is the secondary ecclipse mapping of HD 189733b ($r\sim0.3$) using Spitzer's 8 micron channel \citep{majeau2012} obtained at a precision of about 1 ppt. For a spherical harmonics map of degree $l_{max}=3$ (although $l_{max}=1$ is enough in this case), $q=10$ guarantees a model with a precision of $10^{-10}$, an execution time one order of magnitude faster than \textsf{starry} on CPU and up to two orders of magnitude faster on GPU for large datasets (see right plot of \autoref{real_benchmark}).\\\\
\begin{figure}[H]
    \begin{center}
        \includegraphics[width=\textwidth]{../workflows/speed/figures/speed_vs_N_realistic.pdf}
        \caption{}
        \label{fig:real_benchmark}
    \end{center}
\end{figure}
